cardinalR: Generating interesting high-dimensional data structures

### 2025-09-23 paper-cardinalR.Rmd
Updated HTML and PDF Figures

```{r}
linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
expbranches_lang <- langevitour::langevitour(expbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=15, fig.height=15}

@@

```{r}
gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
quadratic_lang <- langevitour::langevitour(quadratic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
cubic_lang <- langevitour::langevitour(cubic, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
scurvehole_lang <- langevitour::langevitour(scurvehole, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
crescent_lang <- langevitour::langevitour(crescent, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
helicalspiral_lang <- langevitour::langevitour(helicalspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`). The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data.

@@

```{r}
trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

@@

```{r}
trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "500px", height = "400px")
```

>>>>

```{r}
linearbranches_lang <- langevitour::langevitour(linearbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
curvybranches_lang <- langevitour::langevitour(curvybranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
expbranches_lang <- langevitour::langevitour(expbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
orglinearbranches_lang <- langevitour::langevitour(orglinearbranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
orgcurvybranches_lang <- langevitour::langevitour(orgcurvybranches, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
cone_lang <- langevitour::langevitour(cone, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
gridcube_lang <- langevitour::langevitour(gridcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
unifcube_lang <- langevitour::langevitour(unifcube, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
cubehole_lang <- langevitour::langevitour(cubehole, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r cube-proj, eval=knitr::is_latex_output(), fig.cap="Three $2\\text{-}D$ projections from $4\\text{-}D$, for the `gridcube` (a1-a3), `unifcube` (b1-b3), and `cubehole` (c1-c3) data.", fig.width=12, fig.height=15}

@@

```{r}
gau_lang <- langevitour::langevitour(gau, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
linear_lang <- langevitour::langevitour(linear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
mobius_lang <- langevitour::langevitour(mobius, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
quadratic_lang <- langevitour::langevitour(quadratic, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
cubic_lang <- langevitour::langevitour(cubic, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
pyrrect_lang <- langevitour::langevitour(pyrrect, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
pyrtri_lang <- langevitour::langevitour(pyrtri, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
pyrstar_lang <- langevitour::langevitour(pyrstar, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
pyrholes_lang <- langevitour::langevitour(pyrholes, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
scurve_lang <- langevitour::langevitour(scurve, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
scurvehole_lang <- langevitour::langevitour(scurvehole, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
circle_lang <- langevitour::langevitour(circle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
curvycycle_lang <- langevitour::langevitour(curvycycle, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
unifsphere_lang <- langevitour::langevitour(unifsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
gridedsphere_lang <- langevitour::langevitour(gridedsphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
clusteredspheres_lang <- langevitour::langevitour(clusteredspheres, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
hemisphere_lang <- langevitour::langevitour(hemisphere, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
swissroll_lang <- langevitour::langevitour(swissroll, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
crescent_lang <- langevitour::langevitour(crescent, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
curvycylinder_lang <- langevitour::langevitour(curvycylinder, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
sphericalspiral_lang <- langevitour::langevitour(sphericalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
helicalspiral_lang <- langevitour::langevitour(helicalspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
conicspiral_lang <- langevitour::langevitour(conicspiral, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
nonlinear_lang <- langevitour::langevitour(nonlinear, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

The Trefoil is a closed, nontrivial one-dimensional manifold embedded in $3\text{-}D$ or $4\text{-}D$ space (`r knitr::asis_output(if (knitr::is_html_output()) { "Figure \\@ref(fig:trefoil)"} else { "Figure \\@ref(fig:trefoil-proj)"})`). The trefoil features topological complexity in the form of self-overlaps, making it a valuable test case for evaluating the ability of non-linear dimension reduction methods to preserve global structure, loops, and embeddings in high-dimensional data. Table `r knitr::asis_output(ifelse(knitr::is_html_output(), '\\@ref(tab:trefoil-tb-html)', '\\@ref(tab:trefoil-tb-pdf)'))` summarizes these functions.

```{r}
trefoil_tb <- tibble(fun = c("gen_trefoil4d",
                            "gen_trefoil3d"), 
                        exp = c("Trefoil in $4\\text{-}D$.",
                                "Trefoil in $3\\text{-}D$."))
```

```{r trefoil-tb-html, eval=knitr::is_html_output()}
trefoil_tb |> 
  kable(caption = "cardinalR trefoil data generation functions", col.names = c("Function", "Explanation")) 
```

```{r trefoil-tb-pdf, eval=knitr::is_latex_output()}

trefoil_tb |> 
  kable(caption = "cardinalR trefoil data generation functions", format="latex", col.names = c("Function", "Explanation"), booktabs = T)  |>
  column_spec(1, width = "4cm") |>
  column_spec(2, width = "8cm")
```

@@

```{r}
trefoil4d_lang <- langevitour::langevitour(trefoil4d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```

@@

```{r}
trefoil3d_lang <- langevitour::langevitour(trefoil3d, levelColors = "black", enableControls = FALSE, width = "400px", height = "300px")
```
